---
title: "Mini Data Analysis Milestone 2"
output: github_document
---

*To complete this milestone, you can either edit [this `.rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the output in the YAML header to `github_document`, before submitting a tagged release on canvas.*

# Welcome to the rest of your mini data analysis project!

In Milestone 1, you explored your data. and came up with research questions. This time, we will finish up our mini data analysis and obtain results for your data by:

-   Making summary tables and graphs
-   Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

We will also explore more in depth the concept of *tidy data.*

**NOTE**: The main purpose of the mini data analysis is to integrate what you learn in class in an analysis. Although each milestone provides a framework for you to conduct your analysis, it's possible that you might find the instructions too rigid for your data set. If this is the case, you may deviate from the instructions -- just make sure you're demonstrating a wide range of tools and techniques taught in this class.

# Instructions

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-2.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work here--->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to your mini-analysis GitHub repository, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 50 points: 45 for your analysis, and 5 for overall reproducibility, cleanliness, and coherence of the Github submission.

**Research Questions**: In Milestone 1, you chose two research questions to focus on. Wherever realistic, your work in this milestone should relate to these research questions whenever we ask for justification behind your work. In the case that some tasks in this milestone don't align well with one of your research questions, feel free to discuss your results in the context of a different research question.

# Learning Objectives

By the end of this milestone, you should:

-   Understand what *tidy* data is, and how to create it using `tidyr`.
-   Generate a reproducible and clear report using R Markdown.
-   Manipulating special data types in R: factors and/or dates and times.
-   Fitting a model object to your data, and extract a result.
-   Reading and writing data as separate files.

# Setup

Begin by loading your data and the tidyverse package below:

```{r, message = FALSE}
library(datateachr) # <- might contain the data you picked!
library(tidyverse)
library(dplyr)
library(ggplot2)

```

# Task 1: Process and summarize your data

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various data manipulation functions.

### 1.1 (1 point)

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->

1.  Is there a correlation between the year a building was built (year_built) and the availability of energy-efficient features such as separate gas/hydro meters, sprinkler systems, or thermal windows?
2.  How does the number of accessible parking spaces relate to the size of the building (number of storeys or number of units)?
3.  What is the distribution of barrier-free accessibility features in the dataset, and how does it vary across different types of buildings (e.g., privately owned vs. social housing)?
4.  What is the relationship between the number of barrier-free accessible units and the number of elevators in a building?\
    <!----------------------------------------------------------------------------->

Here, we will investigate your data using various data manipulation and graphing functions.

### 1.2 (8 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question.

Ensure that the output of each operation is printed!

Also make sure that you're using dplyr and ggplot2 rather than base R. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

**Summarizing:**

1.  Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2.  Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3.  Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4.  Compute the proportion and counts in each category of one categorical variable across the groups of another categorical variable from your data. Do not use the function `table()`!

**Graphing:**

6.  Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read.
7.  Make a graph where it makes sense to customize the alpha transparency.

Using variables and/or tables you made in one of the "Summarizing" tasks:

8.  Create a graph that has at least two geom layers.
9.  Create 3 histograms, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->

#### 1.2.1 Is there a correlation between the year a building was built (year_built) and the availability of energy-efficient features such as separate gas/hydro meters, sprinkler systems, or thermal windows?

```{r}

# Compute the range, mean, and two other summary statistics of the year built across the groups of a categorical variable.

# Create a categorical variable based on the year_built
buildings_year_stat <- apt_buildings %>%
  mutate(period = case_when(
    year_built < 1950 ~ "Before 1950",
    between(year_built, 1950, 1970) ~ "1950-1970",
    between(year_built, 1970, 1990) ~ "1970-1990",
    between(year_built, 1990, 2010) ~ "1990-2010",
    year_built > 2010 ~ "After 2010"
  ))

# Compute summary statistics
ans1.2.1 <- buildings_year_stat %>%
  group_by(period) %>%
  summarize(
    range = paste(min(year_built), "-", max(year_built)),
    mean_year = mean(year_built),
    median_year = median(year_built),
    sd_year = sd(year_built)
  )

print(ans1.2.1)

```

This task allows us to understand the distribution of building years over different periods.

```{r}
# Make a graph where it makes sense to customize the alpha transparency.

# Create a scatter plot to visualize the relationship
ggplot(buildings_year_stat, aes(x = year_built, y = separate_gas_meters, color = period)) +
  geom_point(alpha = 0.7) +
  labs(title = "Year Built vs. Separate Gas Meters",
       x = "Year Built",
       y = "Separate Gas Meters") +
  theme_minimal()

```

Scatter plots are suitable for visualizing the relationship between the year of construction of a building and the availability of individual gas meters. Use alpha transparency to help visualize overlapping points and understand the density of data points. Different colors represent different periods, providing a comprehensive view of correlations.

#### 1.2.2 How does the number of accessible parking spaces relate to the size of the building (number of storeys or number of units)?

```{r}
# Create a categorical variable with 3 or more groups from an existing numerical variable.

# Define thresholds for categorization
threshold_low <- 10
threshold_high <- 30

# Create a categorical variable
ans1.2.2 <- apt_buildings %>%
  mutate(parking_category = case_when(
    no_of_accessible_parking_spaces <= threshold_low ~ "Few",
    between(no_of_accessible_parking_spaces, threshold_low + 1, threshold_high) ~ "Moderate",
    no_of_accessible_parking_spaces > threshold_high ~ "Many"
  ))

# Display the updated dataset
head(ans1.2.2)

```

By classifying the number of available parking spaces, we can better understand the distribution and make comparisons between different groups.

```{r}
# Create a graph that has at least two geom layers.

ggplot(ans1.2.2, aes(x = factor(no_of_storeys), fill = parking_category)) +
  geom_bar(position = position_dodge2(preserve = "single"), color = "white", alpha = 0.8, width = 0.8) +
  labs(title = "Accessible Parking Spaces vs. Building Size",
       x = "Number of Storeys",
       y = "Count") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = -30, hjust = 0)
  )

```

This bar chart visually displays the relationship between the number of parking spaces available and the size of the building (number of floors). I use different colors for each parking category allows for clear comparisons.

#### 1.2.3 What is the distribution of barrier-free accessibility features in the dataset, and how does it vary across different types of buildings (e.g., privately owned vs. social housing)?

```{r}
# Compute the proportion and counts in each category of one categorical variable across the groups of another categorical variable from your data.

# Compute the counts and proportions
ans1.2.3 <- apt_buildings %>%
  group_by(property_type, barrier_free_accessibilty_entr) %>%
  summarise(count = n()) %>%
  mutate(proportion = count / sum(count))

ans1.2.3
```

This task provides insights into the distribution of accessibility features in different types of buildings. By counting quantities and proportions, we can understand how common these features are across various building categories, helping to address research questions about variations in accessibility between different building types.

```{r}
# Create a stacked bar plot with customized alpha transparency

ggplot(ans1.2.3, aes(x = property_type, y = proportion, fill = barrier_free_accessibilty_entr)) +
  geom_bar(stat = "identity", position = "stack", alpha = 0.8) +
  labs(title = "Distribution of Barrier-Free Accessibility Features",
       x = "Property Type",
       y = "Proportion") +
  theme_minimal()
```

A stacked bar plot is suitable for visualizing the distribution of barrier-free accessibility features across different types of buildings.

#### 1.2.4 What is the relationship between the number of barrier-free accessible units and the number of elevators in a building?

```{r}
# Compute the range, mean, and two other summary statistics of one numerical variable across the groups of one categorical variable from your data.

ans1.2.4 <- apt_buildings %>%
  group_by(no_of_elevators) %>%
  summarize(
    mean_barrier_units = mean(no_barrier_free_accessible_units),
    range_barrier_units = max(no_barrier_free_accessible_units) - min(no_barrier_free_accessible_units),
    sd_barrier_units = sd(no_barrier_free_accessible_units),
    median_barrier_units = median(no_barrier_free_accessible_units)
  )

print(ans1.2.4)
```

This task provides a statistical summary of the relationship between the number of accessible units and the number of elevators in a building. Means, ranges, and other statistics help understand the central tendency and variability of accessible units across elevator categories.

```{r}
apt_buildings %>%
  ggplot(aes(x = no_of_elevators, y = no_barrier_free_accessible_units)) +
  geom_point() +
  labs(
    title = "Relationship Between Barrier-Free Accessible Units and Number of Elevators",
    x = "Number of Elevators",
    y = "Number of Barrier-Free Accessible Units"
  )
```

A scatter plot is chosen to visualize the relationship between two numerical variables, the number of barrier-free accessible units, and the number of elevators.

<!----------------------------------------------------------------------------->

### 1.3 (2 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->

1.  Is there a correlation between the year a building was built (year_built) and the availability of energy-efficient features such as separate gas/hydro meters, sprinkler systems, or thermal windows?

    In this problem we use a scatter plot to illustrate the relationship between construction time and separate gas meter. We can clearly see the difference between having a separate gas meter and not having a separate gas meter in different time periods. I think this question is answered very well and does not need to be modified.

2.  How does the number of accessible parking spaces relate to the size of the building (number of storeys or number of units)?

    In this problem, we can see the distribution of parking spaces on different floors through histograms. We classify parking spaces to make the presentation of the histogram more intuitive. I think this question is answered very well and does not need to be modified.

3.  What is the distribution of barrier-free accessibility features in the dataset, and how does it vary across different types of buildings (e.g., privately owned vs. social housing)?

    In this question, we use a histogram and can clearly see the proportion of different accessibility facilities in different property types. I think this question is answered very well and does not need to be modified.

4.  What is the relationship between the number of barrier-free accessible units and the number of elevators in a building?

    In the answer to this question, it is difficult to observe the relationship between barrier-free facilities and the number of elevators through a scatter plot. I think this question can be modified to study the relationship between barrier-free facilities and building floors

<!----------------------------------------------------------------------------->

# Task 2: Tidy your data

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

-   Each row is an **observation**
-   Each column is a **variable**
-   Each cell is a **value**

### 2.1 (2 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have \>8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->

I select 8 variables in the apt_buildings dataset. From the selected dataset ans2.1 shown below, we can see that each column represents a variable, each row an observation, and each cell a value. The dataset is already a tidy dataset.

```{r}
ans2.1 <- apt_buildings %>% 
  select(id, no_of_elevators, rsn, year_built, year_registered, no_of_storeys, no_of_units, no_barrier_free_accessible_units)
print(ans2.1)
```

<!----------------------------------------------------------------------------->

### 2.2 (4 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->

```{r}
# tidy to untidy
untidy_2.2 <- ans2.1 %>%
  gather(key = "variable", value = "value", -id) %>%
  arrange(id)

untidy_2.2
```

```{r}
# untidy to tidy
tidy_2.2 <- untidy_2.2 %>%
  spread(key = "variable", value = "value")

tidy_2.2
```

<!----------------------------------------------------------------------------->

### 2.3 (4 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in the remaining tasks:

<!-------------------------- Start your work below ---------------------------->

1.  *Is there a correlation between the year a building was built (year_built) and the availability of energy-efficient features such as separate gas/hydro meters, sprinkler systems, or thermal windows?*
2.  *How does the number of accessible parking spaces relate to the size of the building (number of storeys or number of units)?*

<!----------------------------------------------------------------------------->

Explain your decision for choosing the above two research questions.

<!--------------------------- Start your work below --------------------------->

1.  This question is valuable for understanding the evolution of energy-efficient features in buildings over time. By examining the correlation between the year a building was built and the availability of energy-efficient features, we can identify trends and patterns in the adoption of these technologies.
2.  Investigating how the number of accessible parking spaces relates to the size of the building is crucial for urban planning and accessibility considerations. Understanding whether larger buildings tend to provide more accessible parking spaces is essential for designing inclusive and convenient living spaces.

<!----------------------------------------------------------------------------->

Now, try to choose a version of your data that you think will be appropriate to answer these 2 questions. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelevant columns, etc.).

(If it makes more sense, then you can make/pick two versions of your data, one for each research question.)

<!--------------------------- Start your work below --------------------------->

```{r}

data_2.3.1 <- apt_buildings %>%
  # Filtering relevant columns
  select(year_built, separate_gas_meters, separate_hydro_meters, sprinkler_system) %>%
  
  # Handling missing values
  na.omit()%>%
  # create new variable
  mutate(year_since_built = 2023 - year_built)%>%
  select(year_since_built, everything())%>%
  # arrange based on year_since_built
  arrange(year_since_built)

data_2.3.1
```

```{r}
# Filtering relevant columns
data_2.3.2 <- apt_buildings %>%
  select(no_of_accessible_parking_spaces, no_of_storeys, no_of_units) %>%
  
  # Cleaning: Handling missing values
  na.omit() %>%
  
  # Tidying: Renaming columns for better clarity
  rename(
    accessible_parking_spaces = no_of_accessible_parking_spaces,
    num_storeys = no_of_storeys,
    num_units = no_of_units
  ) %>%
  
  # Dropping irrelevant columns
  select(accessible_parking_spaces, num_storeys, num_units)%>%
  arrange(accessible_parking_spaces)
data_2.3.2
```

# Task 3: Modelling

## 3.0 (no points)

Pick a research question from 1.2, and pick a variable of interest (we'll call it "Y") that's relevant to the research question. Indicate these.

<!-------------------------- Start your work below ---------------------------->

**Research Question**: How does the number of accessible parking spaces relate to the size of the building (number of storeys)?

**Variable of interest**: The number of accessible parking spaces (accessible_parking_spaces).

<!----------------------------------------------------------------------------->

## 3.1 (3 points)

Fit a model or run a hypothesis test that provides insight on this variable with respect to the research question. Store the model object as a variable, and print its output to screen. We'll omit having to justify your choice, because we don't expect you to know about model specifics in STAT 545.

-   **Note**: It's OK if you don't know how these models/tests work. Here are some examples of things you can do here, but the sky's the limit.

    -   You could fit a model that makes predictions on Y using another variable, by using the `lm()` function.
    -   You could test whether the mean of Y equals 0 using `t.test()`, or maybe the mean across two groups are different using `t.test()`, or maybe the mean across multiple groups are different using `anova()` (you may have to pivot your data for the latter two).
    -   You could use `lm()` to test for significance of regression coefficients.

<!-------------------------- Start your work below ---------------------------->

```{r}
# Fit a linear regression model
model <- lm(accessible_parking_spaces ~ num_storeys, data = data_2.3.2)

# Print the model summary
summary(model)
```

<!----------------------------------------------------------------------------->

## 3.2 (3 points)

Produce something relevant from your fitted model: either predictions on Y, or a single value like a regression coefficient or a p-value.

-   Be sure to indicate in writing what you chose to produce.
-   Your code should either output a tibble (in which case you should indicate the column that contains the thing you're looking for), or the thing you're looking for itself.
-   Obtain your results using the `broom` package if possible. If your model is not compatible with the broom function you're needing, then you can obtain your results by some other means, but first indicate which broom function is not compatible.

<!-------------------------- Start your work below ---------------------------->

```{r}
# Load the broom package
library(broom)

# Tidy the model results
tidy_results <- tidy(model)

# Print the tidy results
print(tidy_results)

```

The broom package is a tidy tool for summarizing statistical models. It takes the messy output of built-in modeling functions in R, such as **`lm()`**, and turns them into tidy data frames.

<!----------------------------------------------------------------------------->

# Task 4: Reading and writing data

Get set up for this exercise by making a folder called `output` in the top level of your project folder / repository. You'll be saving things there.

## 4.1 (3 points)

Take a summary table that you made from Task 1, and write it as a csv file in your `output` folder. Use the `here::here()` function.

-   **Robustness criteria**: You should be able to move your Mini Project repository / project folder to some other location on your computer, or move this very Rmd file to another location within your project repository / folder, and your code should still work.
-   **Reproducibility criteria**: You should be able to delete the csv file, and remake it simply by knitting this Rmd file.

<!-------------------------- Start your work below ---------------------------->

```{r}
#install.packages("here") 
library(here)

# Define the path to the output folder using here()
output_path <- here("D:/Users/YRC10/Documents/UBC/MENG2/STAT545/Mini Data-Analysis/mda-YRC10/Milestone2")

# Define the file path for the CSV file
csv_file_path <- here(output_path, "Research_Q2.csv")

# Write the summary table to a CSV file
write_csv(data_2.3.2, csv_file_path)

```

<!----------------------------------------------------------------------------->

## 4.2 (3 points)

Write your model object from Task 3 to an R binary file (an RDS), and load it again. Be sure to save the binary file in your `output` folder. Use the functions `saveRDS()` and `readRDS()`.

-   The same robustness and reproducibility criteria as in 4.1 apply here.

<!-------------------------- Start your work below ---------------------------->

```{r}
# Define the file path for the RDS file
rds_file_path <- here(output_path, "linear_model_task3.rds")

# Save the model object to an RDS file
saveRDS(model, rds_file_path)

# Load the model object from the RDS file
loaded_model <- readRDS(rds_file_path)
loaded_model
```

<!----------------------------------------------------------------------------->

# Overall Reproducibility/Cleanliness/Coherence Checklist

Here are the criteria we're looking for.

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors.

The README file should still satisfy the criteria from the last milestone, i.e. it has been updated to match the changes to the repository made in this milestone.

## File and folder structure (1 points)

You should have at least three folders in the top level of your repository: one for each milestone, and one output folder. If there are any other folders, these are explained in the main README.

Each milestone document is contained in its respective folder, and nowhere else.

Every level-1 folder (that is, the ones stored in the top level, like "Milestone1" and "output") has a `README` file, explaining in a sentence or two what is in the folder, in plain language (it's enough to say something like "This folder contains the source for Milestone 1").

## Output (1 point)

All output is recent and relevant:

-   All Rmd files have been `knit`ted to their output md files.
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

Our recommendation: delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant.

## Tagged release (0.5 point)

You've tagged a release for Milestone 2.

### Attribution

Thanks to Victor Yuan for mostly putting this together.
